---
title: "roosevelt_marcellus_wells"
author: "Will Schrepferman"
date: "6/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, include = TRUE)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(janitor)
library(gt)
```

# Introduction

All data from: http://www.depreportingservices.state.pa.us/ReportServer/Pages/ReportViewer.aspx?%2fOil_Gas%2fOil_Gas_Well_Production

I downloaded data for unconventional (shale) wells in the region at the beginning of every year from 2015 through 2020 (the data was formatted differently in previous years and wouldn't be comparable), and filtered for our relevant counties: Allegheny, Armstrong, Beaver, Butler, Cambria, Fayette, Greene, Indiana, Lawrence, Mercer, Somerset, Washington, and Westmoreland.

```{r read_data, include = FALSE}
raw_data <- read.csv("oil_gas_wells.csv") %>%
  clean_names()


# filter for relevant data

clean_data <- raw_data %>%
  mutate(year = as.numeric(paste("20", substring(period_id, 1, 2), sep = ""))) %>%
  mutate(permit_num = as.character(i_well_permit_num)) %>%
  mutate(farm_name_well = as.character(farm_name_well_num)) %>%
  mutate(spud = as.Date(spud_date)) %>%
  mutate(gas_prod_monthly_quantity = as.numeric(gas_quantity)) %>%
  mutate(operating_firm = as.character(operator)) %>%
  mutate(county = as.character(well_county)) %>%
  mutate(municipality = as.character(well_municipality)) %>%
  mutate(latitude = well_latitude) %>%
  mutate(longitude = well_longitude) %>%
  select(year, permit_num, farm_name_well, spud, gas_prod_monthly_quantity, operating_firm, county, municipality, latitude, longitude)


# Counties: Allegheny,Armstrong,Beaver,Butler,Cambria,Fayette,Greene,Indiana,Lawrence,Mercer,Somerset,Washington,and Westmoreland
# from Summer Research Memo


counties_data <- clean_data %>%
  filter(county == "ALLEGHENY" | county == "ARMSTRONG" | county == "BEAVER" | county == "BUTLER" | county == "CAMBRIA" | county == "FAYETTE" | 
         county == "GREENE" | county == "INDIANA" | county == "LAWRENCE" | county == "MERCER" | county == "SOMERSET" | county == "WASHINGTON" | 
         county == "WESTMORELAND")


```


Since 2015, the total number of wells in the Marcellus Shale within these specified counties has been increasing, but that increase appears to be slowing.

```{r graph1}
counties_data %>%
  group_by(year) %>%
  summarize(count = n()) %>%
  ggplot(aes(x = year, y = count)) +
  geom_bar(stat = "identity") +
  labs(title = "Shale Wells in the Marcellus Region since 2015", subtitle = "Relevant Counties")
```



### Here are the top 10 firms in every given year:

```{r graph2}
graph_2015 <- counties_data %>%
  filter(year == 2015) %>%
  group_by(operating_firm) %>%
  summarize(count = n()) %>%
  arrange(-count) %>%
  head(10) %>%
  ggplot(aes(x = reorder(operating_firm, -count), y = count)) +
  geom_text(aes(label=count), vjust=-0.15, size=3.5) +
  theme(axis.text.x = element_text(angle = 90, size = 7.5, hjust = 0.5)) +
  geom_bar(stat = "identity") + 
  ylim(0, 1400) +
  labs(title = "Wells in 2015 (Top 10 Firms)", x = "Firm", y = "No. of Wells")

graph_2016 <- counties_data %>%
  filter(year == 2016) %>%
  group_by(operating_firm) %>%
  summarize(count = n()) %>%
  arrange(-count) %>%
  head(10) %>%
  ggplot(aes(x = reorder(operating_firm, -count), y = count)) +
  geom_text(aes(label=count), vjust=-0.15, size=3.5) +
  theme(axis.text.x = element_text(angle = 90, size = 7.5, hjust = 0.5)) +
  geom_bar(stat = "identity") + 
  ylim(0, 1400) +
  labs(title = "Wells in 2016 (Top 10 Firms)", x = "Firm", y = "No. of Wells")

graph_2017 <- counties_data %>%
  filter(year == 2017) %>%
  group_by(operating_firm) %>%
  summarize(count = n()) %>%
  arrange(-count) %>%
  head(10) %>%
  ggplot(aes(x = reorder(operating_firm, -count), y = count)) +
  geom_text(aes(label=count), vjust=-0.15, size=3.5) +
  theme(axis.text.x = element_text(angle = 90, size = 7.5, hjust = 0.5)) +
  geom_bar(stat = "identity") + 
  ylim(0, 1400) +
  labs(title = "Wells in 2017 (Top 10 Firms)", x = "Firm", y = "No. of Wells")

graph_2018 <- counties_data %>%
  filter(year == 2018) %>%
  group_by(operating_firm) %>%
  summarize(count = n()) %>%
  arrange(-count) %>%
  head(10) %>%
  ggplot(aes(x = reorder(operating_firm, -count), y = count)) +
  geom_text(aes(label=count), vjust=-0.15, size=3.5) +
  theme(axis.text.x = element_text(angle = 90, size = 7.5, hjust = 0.5)) +
  geom_bar(stat = "identity") + 
  ylim(0, 1400) +
  labs(title = "Wells in 2018 (Top 10 Firms)", x = "Firm", y = "No. of Wells")

graph_2019 <- counties_data %>%
  filter(year == 2019) %>%
  group_by(operating_firm) %>%
  summarize(count = n()) %>%
  arrange(-count) %>%
  head(10) %>%
  ggplot(aes(x = reorder(operating_firm, -count), y = count)) +
  geom_text(aes(label=count), vjust=-0.15, size=3.5) +
  theme(axis.text.x = element_text(angle = 90, size = 7.5, hjust = 0.5)) +
  geom_bar(stat = "identity") + 
  ylim(0, 1400) +
  labs(title = "Wells in 2019 (Top 10 Firms)", x = "Firm", y = "No. of Wells")

graph_2020 <- counties_data %>%
  filter(year == 2020) %>%
  group_by(operating_firm) %>%
  summarize(count = n()) %>%
  arrange(-count) %>%
  head(10) %>%
  ggplot(aes(x = reorder(operating_firm, -count), y = count)) +
  geom_text(aes(label=count), vjust=-0.15, size=3.5) +
  theme(axis.text.x = element_text(angle = 90, size = 7.5, hjust = 0.5)) +
  geom_bar(stat = "identity") + 
  ylim(0, 1400) +
  labs(title = "Wells in 2020 (Top 10 Firms)", x = "Firm", y = "No. of Wells")


graph_2015
graph_2016
graph_2017
graph_2018
graph_2019
graph_2020

```



### And here is the data including every firm and well per year:

```{r tables}
plot_2015 <- counties_data %>%
  filter(year == 2015) %>%
  group_by(operating_firm) %>%
  summarize(count = n()) %>%
  arrange(-count) %>%
  gt() %>%
  tab_header(title = "Wells in 2015")

plot_2016 <- counties_data %>%
  filter(year == 2016) %>%
  group_by(operating_firm) %>%
  summarize(count = n()) %>%
  arrange(-count) %>%
  gt() %>%
  tab_header(title = "Wells in 2016")

plot_2017 <- counties_data %>%
  filter(year == 2017) %>%
  group_by(operating_firm) %>%
  summarize(count = n()) %>%
  arrange(-count) %>%
  gt() %>%
  tab_header(title = "Wells in 2017")

plot_2018 <- counties_data %>%
  filter(year == 2018) %>%
  group_by(operating_firm) %>%
  summarize(count = n()) %>%
  arrange(-count) %>%
  gt() %>%
  tab_header(title = "Wells in 2018")

plot_2019 <- counties_data %>%
  filter(year == 2019) %>%
  group_by(operating_firm) %>%
  summarize(count = n()) %>%
  arrange(-count) %>%
  gt() %>%
  tab_header(title = "Wells in 2019")

plot_2020 <- counties_data %>%
  filter(year == 2020) %>%
  group_by(operating_firm) %>%
  summarize(count = n()) %>%
  arrange(-count) %>%
  gt() %>%
  tab_header(title = "Wells in 2020")

plot_2015
plot_2016
plot_2017
plot_2018
plot_2019
plot_2020
```




